Evaluating infectious disease forecasts with allocation scoring rules

### 2023-05-08 alloscore_manuscript/allocation.tex
update section on quantile score and crps

\subsection{A review of quantile loss, CRPS, and the weighted interval score}
\label{sec:methods.quantileloss}

We review here how quantile loss arises from a particular decision-making problem, and how CRPS can be obtained by integrating across values of the parameters of that decision-making problem. These results have been thoroughly discussed in the literature [cite cite cite].

@@

\paragraph{Step 3: Define the scoring rule.} Following the procedures outlined above, we could score the forecast distribution $F$ with the scoring rule
\begin{align}
S_Q(F, y; C, L) &= s_Q(x^F, y; C, L) = Cx^F + L(x^F-y)_- \\
&= C F^{-1}(\alpha) + L \left(F^{-1}(\alpha) - y\right)_-
\end{align}

This scoring rule evaluates the forecast distribution $F$ only through its $\alpha$ quantile. While this is faithful to the context of the decision-making problem, it may not be satisfying as a measure of the quality of the full forecast distribution. For this purpose, one option is to integrate the quantile loss across different values of the quantile level $\alpha$, or, equivalently, across different values of the decision problem's parameters $C$ and $L$. This gives rise to the (weighted) CRPS:
\begin{align*}
    S_{CRPS}(F, y) &= \int S_Q(F, y; C, L) p(C, L) \, dC dL
\end{align*}
The usual CRPS results from taking $p$ to be the density of a distribution such that the induced distribution on the probability level $\alpha = 1 - C/L$ is $\text{Uniform}(0,1)$.
The weighted interval score (WIS) can be used when the full forecast distribution $F$ is not available, and amounts to an approximation of CRPS where the distribution $p$ is discrete with point masses at the quantile levels corresponding to the endpoints of a finite set of prediction intervals.

\elr{maybe better to use different notation for this integral that's less specific about whether $p$ is discrete or continuous?}

% The below was an unsuccessful attempt to connect to a more usual notation for CRPS just in terms of integrating across quantile levels. It resulted in a notational quagmire, but I'm keeping it here for now in case it seems like a better way to go...
%
%To make the connection to the usual notation, we can divide the loss function $s_Q$ by $L$ to obtain an expression in terms of only $\alpha$.
%\begin{align*}
%s_Q(x,y; \alpha) &= s_Q(x,y; C, L)/L \\
%&= (C/L)x + (x-y)_- \\
%&= (1 - \alpha)x + (x-y)_-
%\end{align*}
%Because these loss functions are equal up to a constant of proportionality, the Bayes act is the same for both. We can how obtain a weighted CRPS by integrating the quantile score across values of $C$ and $L$ or across values of $\alpha$:
%\begin{align*}
%S_{CRPS}(F, y) &= \int S_Q(F, y; C, L) h(C, L) \, dC dL %\\
%&= \int S_Q(F, y; \alpha) h(\alpha) \, d\alpha
%\end{align*}
%The usual CRPS results from taking $h$ such that the distribution on the probability level $\alpha$ is $\text{Uniform}(0,1)$.

@@

>>>>

\subsection{A review of quantile score, CRPS, and the weighted interval score}
\label{sec:methods.quantileloss}

We review how the quantile score arises from a particular decision-making problem in section \ref{sec:methods.quantileloss.quantile}, and how CRPS can be obtained by integrating across values of the parameters of that decision-making problem, as well as the connection to WIS, in section \ref{sec:methods.quantileloss.crps}. These results have been thoroughly discussed in the literature [cite cite cite].

\subsubsection{Decision-theoretic origins of the quantile score}
\label{sec:methods.quantileloss.quantile}

@@

\paragraph{Step 3: Define the scoring rule.} Following the procedures outlined above, we could score the forecast distribution $F$ with the scoring rule
\begin{align*}
S_Q(F, y; C, L) &= s_Q(x^F, y; C, L) = Cx^F + L(x^F-y)_- \nonumber \\
&= C F^{-1}(\alpha) + L \left(F^{-1}(\alpha) - y\right)_- \label{eqn:S_Q_CL}
\end{align*}

We have set up the problem here in terms of the cost and loss parameters $C$ and $L$, which has the benefit of an intuitive connection to the decision-making context. However, to clarify the connection to the usual notation for the quantile loss, we can divide the loss function $s_Q$ by $L$ to obtain an expression in terms of only $\alpha$:
\begin{align*}
s_Q(x,y; \alpha) &= s_Q(x,y; C, L)/L \\
&= (C/L)x + (x-y)_- \\
&= (1 - \alpha)x + (x-y)_-.
\end{align*}
Because these loss functions are equal up to a constant of proportionality, the Bayes act is the same for both. The associated quantile scoring rule expressed in terms of $\alpha$ is
\begin{equation}
S_Q(F, y; \alpha) = (1 - \alpha)F^{-1}(\alpha) + \left(F^{-1}(\alpha) - y\right)_-. \label{eqn:S_Q_alpha}
\end{equation}
In either formulation, the key observation is that the Bayes act is the quantile of the forecast distribution $F$ at the probability level given by one minus the cost/loss ratio $C/L$.

\subsubsection{CRPS as an integrated quantile score}
\label{sec:methods.quantileloss.crps}

The scoring rule $S_Q$ of Equation \eqref{eqn:S_Q_alpha} evaluates the forecast distribution $F$ only through its $\alpha$ quantile. While this is faithful to the context of the decision-making problem, it may not be satisfying as a measure of the quality of the full forecast distribution. For this purpose, one option is to integrate the quantile scoring rule across different values of the probability level $\alpha$, weighting the probability levels according to a specified distribution $p$. This yields a weighted CRPS:
\begin{align*}
S_{CRPS}(F, y; p) &= \int S_Q(F, y; \alpha) p(\alpha) \, d\alpha.
\end{align*}
This weighted form of CRPS has appeared in the literature before, e.g. see \textcite{gneiting2011weightedScoringRules}. The usual CRPS results from taking $p$ corresponding to a $\text{Uniform}(0,1)$ distribution, equally weighting all probability levels.

We emphasize that because $\alpha = 1 - C/L$, the distribution $p(\alpha)$ can be interpreted as expressing incomplete knowledge about the cost/loss ratio in the decision-making problem. The equal weighting used by the ordinary CRPS may be appropriate in the absence of any knowledge about the context in which forecasts will be used to support decision-making, but may be inappropriate if more information is known about the cost/loss ratio for a specific decision-making task.

The weighted interval score (WIS) is often used when the full forecast distribution $F$ is not available, as in the U.S. Forecast Hub and similar efforts where forecasts are represented by a collection of prediction intervals. WIS is a discrete approximation to CRPS, and can be obtained by using a distribution $p$ that has point masses at the probability levels corresponding to the endpoints of a finite set of prediction intervals.

%\elr{Leaving the comment below in the tex, but noting that it was in response to an older version of the write up. I've since attempted to address it by moving to a discussion in terms of a version of the scoring function where we've divided by L to work in terms of the cost/loss ratio.}
%\apg{Quick note that I don't think this will work without some more identifications. The integral needs to be over probability levels and thresholds and without the some discussion like the one commented out (currently) at line 123 or (better in my mind) one like I'm trying to do with binary choices at the unit-level, I don't think we have a threshold.}
%The usual CRPS results from taking $p$ to be the density of a distribution such that the induced distribution on the probability level $\alpha = 1 - C/L$ is $\text{Uniform}(0,1)$.

@@

\elr{If we like the organization of the previous subsection into two subsubsections about the quantile score and the CRPS, we should replicate that here.}
